---
title: "Alexander Francisco Lab 3"
output: github_document
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)


load(file.choose())
```

## Check Categories

```{r}
unique(d_HHP2020_24$Region)
unique(d_HHP2020_24$Education)
unique(d_HHP2020_24$Mar_Stat)
```

## Prepare the Data

```{r}


HHP_divorce <- d_HHP2020_24 %>%
  filter(
    !is.na(Region),
    !is.na(Education),
    !is.na(Mar_Stat)
  ) %>%
  filter(
    !str_to_lower(str_trim(as.character(Education))) %in%
      c("lt hs", "some hs")
  ) %>%
  mutate(
    divorced = as.numeric(Mar_Stat == "divorced")
  )

# 
unique(HHP_divorce$Education)
```

## Table 1: Overall Divorce Statistics by Region

```{r}
region_stats <- HHP_divorce %>%
  group_by(Region) %>%
  summarize(
    `Sample Size` = n(),
    `Number Divorced` = sum(divorced),
    `Divorce Rate` = mean(divorced),
    `Standard Deviation` = sd(divorced),
    .groups = "drop"
  )

region_stats_display <- region_stats %>%
  mutate(
    `Sample Size` = format(`Sample Size`, big.mark = ","),
    `Number Divorced` = format(`Number Divorced`, big.mark = ","),
    `Divorce Rate` = paste0(round(`Divorce Rate` * 100, 1), "%"),
    `Standard Deviation` = round(`Standard Deviation`, 3)
  )

kable(
  region_stats_display,
  align = "c",
  caption = "Table 1: Overall Divorce Statistics by Region"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 2: Divorce Statistics by Education Level

```{r}
education_stats <- HHP_divorce %>%
  group_by(Education) %>%
  summarize(
    `Sample Size` = n(),
    `Number Divorced` = sum(divorced),
    `Divorce Rate` = mean(divorced),
    `Standard Deviation` = sd(divorced),
    .groups = "drop"
  )

education_stats_display <- education_stats %>%
  mutate(
    `Sample Size` = format(`Sample Size`, big.mark = ","),
    `Number Divorced` = format(`Number Divorced`, big.mark = ","),
    `Divorce Rate` = paste0(round(`Divorce Rate` * 100, 1), "%"),
    `Standard Deviation` = round(`Standard Deviation`, 3)
  )

kable(
  education_stats_display,
  align = "c",
  caption = "Table 2: Divorce Statistics by Education Level"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 3: Divorce Statistics by Region and Education

```{r}
region_education_stats <- HHP_divorce %>%
  group_by(Region, Education) %>%
  summarize(
    `Sample Size` = n(),
    `Number Divorced` = sum(divorced),
    `Divorce Rate` = mean(divorced),
    `Standard Deviation` = sd(divorced),
    .groups = "drop"
  )

region_education_display <- region_education_stats %>%
  mutate(
    `Sample Size` = format(`Sample Size`, big.mark = ","),
    `Number Divorced` = format(`Number Divorced`, big.mark = ","),
    `Divorce Rate` = paste0(round(`Divorce Rate` * 100, 1), "%"),
    `Standard Deviation` = round(`Standard Deviation`, 3)
  )

kable(
  region_education_display,
  align = "c",
  caption = "Table 3: Divorce Statistics by Region and Education Level"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 4: Divorce Rates by Region and Education

```{r}
divorce_rate_table <- region_education_stats %>%
  select(Region, Education, `Divorce Rate`) %>%
  pivot_wider(
    names_from = Region,
    values_from = `Divorce Rate`
  ) %>%
  mutate(
    across(
      where(is.numeric),
      ~ paste0(round(.x * 100, 1), "%")
    )
  )

kable(
  divorce_rate_table,
  align = "c",
  caption = "Table 4: Divorce Rates Across Regions for Each Education Level"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Function for Regional Comparisons

```{r}
compare_regions <- function(data, region1, region2) {

  temp <- data %>%
    filter(Region %in% c(region1, region2)) %>%
    group_by(Education, Region) %>%
    summarize(
      n = n(),
      rate = mean(divorced),
      .groups = "drop"
    ) %>%
    pivot_wider(
      names_from = Region,
      values_from = c(n, rate)
    )

  n1 <- paste0("n_", region1)
  n2 <- paste0("n_", region2)
  r1 <- paste0("rate_", region1)
  r2 <- paste0("rate_", region2)

  temp %>%
    mutate(
      Difference = .data[[r1]] - .data[[r2]],
      `Standard Error` = sqrt(
        .data[[r1]] * (1 - .data[[r1]]) / .data[[n1]] +
        .data[[r2]] * (1 - .data[[r2]]) / .data[[n2]]
      ),
      `95% CI Lower` = Difference - 1.96 * `Standard Error`,
      `95% CI Upper` = Difference + 1.96 * `Standard Error`
    ) %>%
    transmute(
      Education,
      `Region 1` = region1,
      `Region 1 Rate` = paste0(round(.data[[r1]] * 100, 1), "%"),
      `Region 2` = region2,
      `Region 2 Rate` = paste0(round(.data[[r2]] * 100, 1), "%"),
      `Difference (percentage points)` = round(Difference * 100, 2),
      `Standard Error` = round(`Standard Error`, 4),
      `95% CI Lower` = paste0(round(`95% CI Lower` * 100, 2), "%"),
      `95% CI Upper` = paste0(round(`95% CI Upper` * 100, 2), "%")
    )
}
```

## Table 5: Northeast vs. South

```{r}
NE_South_display <- compare_regions(
  HHP_divorce,
  "Northeast",
  "South"
)

kable(
  NE_South_display,
  align = "c",
  caption = "Table 5: Northeast vs. South Divorce Rates"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 6: Northeast vs. Midwest

```{r}
NE_Midwest_display <- compare_regions(
  HHP_divorce,
  "Northeast",
  "Midwest"
)

kable(
  NE_Midwest_display,
  align = "c",
  caption = "Table 6: Northeast vs. Midwest Divorce Rates"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 7: West vs. South

```{r}
West_South_display <- compare_regions(
  HHP_divorce,
  "West",
  "South"
)

kable(
  West_South_display,
  align = "c",
  caption = "Table 7: West vs. South Divorce Rates"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 8: West vs. Midwest

```{r}
West_Midwest_display <- compare_regions(
  HHP_divorce,
  "West",
  "Midwest"
)

kable(
  West_Midwest_display,
  align = "c",
  caption = "Table 8: West vs. Midwest Divorce Rates"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 9: Highest and Lowest Divorce Rate in Each Region

```{r}
highest_lowest <- region_education_stats %>%
  group_by(Region) %>%
  summarize(
    `Highest Divorce Education` = as.character(
      Education[which.max(`Divorce Rate`)]
    ),
    `Highest Divorce Rate` = max(`Divorce Rate`),
    `Lowest Divorce Education` = as.character(
      Education[which.min(`Divorce Rate`)]
    ),
    `Lowest Divorce Rate` = min(`Divorce Rate`),
    .groups = "drop"
  ) %>%
  mutate(
    `Highest Divorce Rate` =
      paste0(round(`Highest Divorce Rate` * 100, 1), "%"),
    `Lowest Divorce Rate` =
      paste0(round(`Lowest Divorce Rate` * 100, 1), "%")
  )

kable(
  highest_lowest,
  align = "c",
  caption = "Table 9: Highest and Lowest Divorce Rates by Education Within Each Region"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 10: Distribution of Divorced Individuals Across Regions

```{r}
divorced_region_distribution <- HHP_divorce %>%
  filter(divorced == 1) %>%
  count(Region, name = "Number Divorced") %>%
  mutate(
    `Percent of All Divorced` =
      `Number Divorced` / sum(`Number Divorced`)
  )

divorced_region_distribution_display <-
  divorced_region_distribution %>%
  mutate(
    `Number Divorced` =
      format(`Number Divorced`, big.mark = ","),
    `Percent of All Divorced` =
      paste0(round(`Percent of All Divorced` * 100, 1), "%")
  )

kable(
  divorced_region_distribution_display,
  align = "c",
  caption = "Table 10: Distribution of Divorced Individuals Across Regions"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```

## Table 11: Distribution of Divorced Individuals Across Education Levels

```{r}
divorced_education_distribution <- HHP_divorce %>%
  filter(divorced == 1) %>%
  count(Education, name = "Number Divorced") %>%
  mutate(
    `Percent of All Divorced` =
      `Number Divorced` / sum(`Number Divorced`)
  )

divorced_education_distribution_display <-
  divorced_education_distribution %>%
  mutate(
    `Number Divorced` =
      format(`Number Divorced`, big.mark = ","),
    `Percent of All Divorced` =
      paste0(round(`Percent of All Divorced` * 100, 1), "%")
  )

kable(
  divorced_education_distribution_display,
  align = "c",
  caption = "Table 11: Distribution of Divorced Individuals Across Education Levels"
) %>%
  kable_styling(full_width = FALSE, font_size = 14)
```
